Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers

### 2023-04-28 manuscript/manuscript.Rmd
made some cosmetic changes

---
title             : "Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers"
shorttitle        : "Explore and Confirm Analysis Workflow"
author:
  - name: Robert T. Thibault
    affiliation: '1,2'
    corresponding: yes
    email: robert.thibault@stanford.edu
    address: Enter postal address here
  - name: Marton Kovacs
    affiliation: '3,4'
    email: marton.balazs.kovacs@gmail.com
  - name: Tom E. Hardwicke
    affiliation: '5'
  - name: Alexandra Sarafoglou
    affiliation: '6'
  - name: John P. A. Ioannidis
    affiliation: '1,7'
  - name: Marcus R. Munafò
    affiliation: '2,8'
affiliation:
  - id: '1'
    institution: Meta-Research Innovation Center at Stanford (METRICS), Stanford University.
  - id: '2'
    institution: School of Psychological Science, University of Bristol.
  - id: '3'
    institution: Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
  - id: '4'
    institution: Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
  - id: '5'
    institution: Melbourne School of Psychological Sciences, University of Melbourne.
  - id: '6'
    institution: Department of Psychology, University of Amsterdam.
  - id: '7'
    institution: Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for
      Transforming Biomedical Research, Berlin Institute of Health, Charité – Universitätsmedizin
      Berlin.
  - id: '8'
    institution: MRC Integrative Epidemiology Unit at the University of Bristol.
abstract: |
  \textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis. An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW). In this workflow, a data management organization initially provides access to only a subset of their dataset to researchers who request it. The researchers then prepare an analysis script based on the subset of data, upload the analysis script to a registry, and then receive access to the full dataset. ECAWs aim to achieve similar goals as preregistration, but make access to the full dataset contingent on compliance. The present survey aimed to garner information from the research community where ECAWs could be applied—employing the Avon Longitudinal Study of Parents and Children (ALSPAC) as a case example. \hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC’s transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}. We received 107 responses, for a 10% response rate. The results suggest that—at least among our sample of respondents—ECAWs hold the potential to serve their intended purpose and appear relatively acceptable. For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55% who agreed). However, as many as 25% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions. \hfill \break \hfill \break \textbf{Preregistration:} https://osf.io/g2fw5 Deviations from the preregistration are outlined in Supplementary Material A.

@@

classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
    extra_dependencies: "tcolorbox"
header-includes   :
  - \usepackage{tcolorbox}
  - \usepackage{caption}
  - \captionsetup[figure]{labelformat=empty}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
# Used for rendering the manuscript rmarkdown document
library("papaja")
# Used for general data wrangling
library(tidyverse)
# Used to make exact file paths
library(here)
# Used for tables
library(kableExtra)
library(gt)
# Used for joining figures
library(patchwork)
# The developer version is needed
# Used for creating Likert ggplots
library(likert)
# Used for knitting rmarkdown documents
library(knitr)
# Used for colorblind ggplot colors
library(viridis)
# References
r_refs("ecaw-ref.bib")
```

@@

\begin{tcolorbox}
\textbf{Box 1. A hypothetical example of an ECAW in practice.}

\includegraphics[width=1\textwidth]{./ecaw_workflow.png}
{\fontsize{10pt}{10pt}\selectfont
The overarching framework for ECAWs is depicted above. Within this framework, data management organizations would need to pre-specify several details for each step. Below, we provide a hypothetical example for illustrative purposes. The exact implementation of ECAWs would depend on the preferences of each data management organization and the community who uses their dataset.

@@

# Results
## Survey completion
Of 1148 emails sent, 1094 went through and 54 bounced. The survey was completed `r nrow(processed)` times and partially completed `r nrow(processed_all) - nrow(processed)` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(processed_all) / 1094 * 100)`% for at least partially complete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to. The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action. Thus, their list likely includes several email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years. It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions). These two factors may have deflated the response rate.} The median time to complete the survey was `r round(median(processed$duration_in_mins), 1)` minutes (IQR: `r round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1)` to `r round(quantile(processed$duration_in_mins, .75, na.rm = T), 1)`). This manuscript presents the results for complete surveys. Supplementary Material D presents the results with partially complete surveys included.

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 2, top panel). At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).

```{r typicallyEcawPlot, warning=FALSE, fig.cap="(ref:typicallyEcawPlotCaption)", fig.align="center", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
# Create typically trustworthy and reproducible plot
# Check if there is any missing and don't understand responses separately
# processed %>%
#   select(contains("typically_")) %>%
#   dplyr::filter(is.na(typically_reproducible) | is.na(typically_trustworthy) | typically_reproducible == "I don't understand the question" | typically_trustworthy == "I don't understand the question") %>%
#   nrow()

# Prepare plot data
typically_plot_data <-
  processed %>%
  select(contains("typically_")) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Strongly disagree",
        "Somewhat disagree",
        "Neither agree nor disagree",
        "Somewhat agree",
        "Strongly agree"
      ))
    )
  ) %>%
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

@@

(ref:typicallyEcawPlotCaption) __Figure 1. Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust”. The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.” For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_. The bar charts in the bottom panel excluded responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible"), n)`).

@@

```{r methodPlot, warning=FALSE, fig.cap="(ref:methodPlotCaption)", fig.align="center", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
# Used methods figures
method_plot_data <-
  processed %>%
  select(starts_with("method_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "scale",
  ) %>%
  separate(variable, into = c("name_prefix", "type")) %>%
  group_by(type) %>%
  count(scale) %>%
  ungroup() %>%
  replace_na(list(scale = "Missing")) %>%
  mutate(
    scale = factor(scale, levels = c(
      "Never or almost never",
      "Sometimes",
      "About half the time",
      "Most of the time",
      "Always or almost always",
      "Missing",
      "I don't understand the question"
    )),
    type = case_when(
      type == "blind" ~ "Blind the data analyst",
      type == "confirmatory" ~ "Contain confirmatory analysis",
      type == "exploratory" ~ "Contain exploratory analysis",
      type == "preregistered" ~ "Are preregistered",
      type == "script" ~ "Share analysis scripts",
    ),
    type = factor(type, levels = c(
      "Contain exploratory analysis",
      "Contain confirmatory analysis",
      "Blind the data analyst",
      "Share analysis scripts",
      "Are preregistered"
    ))
  ) %>%
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>%
  mutate(
    percentage = n / sum(n)
  ) %>%
  ungroup()

@@

(ref:methodPlotCaption) __Figure 2. Responses to survey questions about the research practices of participants.__

@@

```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotCaption)", fig.align="center", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
# Check if there is any missing and don't understand responses separately
alspac_missing <-
  processed %>%
  select(contains("alspac_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
  ) %>%
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>%
  mutate(
    item = str_to_title(str_remove(item, "^[^_]*_")),
    # We reorder it by making it a factor
    item = factor(item, levels = c(
      "Less_willing",
      "Opt_in",
      "Study",
      "Prefer_ecaw"
    )),
    # By making it a factor we can complete missing levels in the data later
    value = factor(value, levels = c(
      "Missing",
      "I don't understand the question",
      "Unsure"
    ))
  ) %>%
  replace_na(list(value = "Missing")) %>%
  group_by(item, value) %>%
  summarise(
    n = n()
  ) %>% 
  # Add n = 0 if no one choose a specific factor level
  tidyr::complete(value, fill = list(n = 0))

@@

(ref:alspacPlotCaption) __Figure 3. Responses to survey questions about using ECAWs.__ These bar charts exclude responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`; respectively from top to bottom), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`). Agreement with the first question may be slightly inflated due to the format of the questions in this block. Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions. Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question \protect\footnotemark[3]. Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[4].

@@

__Table 1. Recurring topics in responses to the open-ended survey questions.__ The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments. These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents. A complete list of responses are viewable in the open data [LINK]. We synthesized the responses to open-ended questions into the 9 topics on the left side of this table. We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as conventional preregistration and ECAWs. On the right side of the table, we provide our reflection on each topic.

```{r openEndedTable, echo=FALSE, message=FALSE, warnings=FALSE, fig.align = "center", out.width="100%", fig.id=FALSE}

@@

open_ended_table %>% 
  group_by(subheader) %>% 
  gt::gt(.) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_row_groups(), cells_column_labels())
    ) %>% 
  tab_footnote(
    footnote = "The text in the Respondents’ comments column is synthesized from multiple responses and/or paraphrased. They are not verbatim responses.",
    locations = cells_column_labels(
      columns = `Respondents’ comments`
    )
  ) %>% 
  tab_footnote(
    footnote = "Sarafoglou et al.’s study used a pre-cleaned dataset. We are not aware of data that elucidates the time difference between ECAW-like workflows and conventional preregistration when starting from raw data.",
    locations = cells_body(
      columns = `Our reflection`,
      rows = 1
      )
  ) %>% 
  tab_options(table.width = pct(90)) %>% 
  gtsave("figs/open-ended-table_figure.png", expand = c(0, 100, 0, 100), zoom = 2, vwidth = 1300)

@@

## Exploratory analyses
Results can be explored interactively at [TEMPORARY LINK], or by running the code available at [INSERT DATA LINK]. Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.

@@

## Contributions
__Conceptualization:__ Robert T. Thibault and Marcus R. Munafò.
__Data curation:__ Marton Kovacs.
__Formal analysis:__ Robert T. Thibault and Marton Kovacs.
__Funding acquisition:__ Robert T. Thibault and John P. A. Ioannidis.
__Investigation:__ Robert T. Thibault.
__Methodology:__ Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.
__Project administration:__ Robert T. Thibault.
__Resources:__ Robert T. Thibault.
__Software:__ Marton Kovacs.
__Supervision:__ Robert T. Thibault and Marcus R. Munafò.
__Validation:__ Robert T. Thibault and Marton Kovacs.
__Visualization:__ Robert T. Thibault and Marton Kovacs.
__Writing - original draft:__ Robert T. Thibault.
__Writing - review & editing:__ Robert T. Thibault, Marton Kovacs, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.

@@

# (APPENDIX) Appendix {-}

>>>>

---
title             : "Reducing bias in secondary data analysis via an Explore and Confirm Analysis Workflow (ECAW): A proposal and survey of observational researchers"
shorttitle        : "Explore and Confirm Analysis Workflow"
author:
  - name: Robert T. Thibault
    affiliation: '1,2'
    corresponding: yes
    email: robert.thibault@stanford.edu
  - name: Marton Kovacs
    affiliation: '3,4'
    email: marton.balazs.kovacs@gmail.com
  - name: Tom E. Hardwicke
    affiliation: '5'
  - name: Alexandra Sarafoglou
    affiliation: '6'
  - name: John P. A. Ioannidis
    affiliation: '1,7'
  - name: Marcus R. Munafò
    affiliation: '2,8'
affiliation:
  - id: '1'
    institution: Meta-Research Innovation Center at Stanford (METRICS), Stanford University.
  - id: '2'
    institution: School of Psychological Science, University of Bristol.
  - id: '3'
    institution: Doctoral School of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
  - id: '4'
    institution: Institute of Psychology, ELTE Eotvos Lorand University, Budapest, Hungary
  - id: '5'
    institution: Melbourne School of Psychological Sciences, University of Melbourne.
  - id: '6'
    institution: Department of Psychology, University of Amsterdam.
  - id: '7'
    institution: Meta-Research Innovation Center Berlin (METRIC-B), QUEST Center for
      Transforming Biomedical Research, Berlin Institute of Health, Charité – Universitätsmedizin
      Berlin.
  - id: '8'
    institution: MRC Integrative Epidemiology Unit at the University of Bristol.
abstract: |
  \justifying\textbf{Background.} Although preregistration can reduce researcher bias and increase transparency in primary research settings, it is less applicable to secondary data analysis. An alternative method that affords additional protection from researcher bias, which cannot be gained from conventional forms of preregistration alone, is an Explore and Confirm Analysis Workflow (ECAW). In this workflow, a data management organization initially provides access to only a subset of their dataset to researchers who request it. The researchers then prepare an analysis script based on the subset of data, upload the analysis script to a registry, and then receive access to the full dataset. ECAWs aim to achieve similar goals as preregistration, but make access to the full dataset contingent on compliance. The present survey aimed to garner information from the research community where ECAWs could be applied—employing the Avon Longitudinal Study of Parents and Children (ALSPAC) as a case example. \hfill \break \hfill \break \textbf{Methods.} We emailed a web-based survey to researchers who had previously applied for access to ALSPAC’s transgenerational observational dataset.\hfill \break \hfill \break \textbf{Results}. We received 107 responses, for a 10% response rate. The results suggest that—at least among our sample of respondents—ECAWs hold the potential to serve their intended purpose and appear relatively acceptable. For example, only 10% of respondents disagreed that ALSPAC should run a study on ECAWs (versus 55% who agreed). However, as many as 25% of respondents agreed that they would be less willing to use ALSPAC data if they were required to use an ECAW (versus 44% who disagreed).\hfill \break \hfill \break \textbf{Conclusion.} Our data and findings provide information for organizations and individuals interested in implementing ECAWs and related interventions. \hfill \break \hfill \break \textbf{Preregistration:} https://osf.io/g2fw5 Deviations from the preregistration are outlined in Supplementary Material A.

@@

classoption       : "man"
output            :
  papaja::apa6_pdf:
    latex_engine: xelatex
    extra_dependencies: ["tcolorbox", "relsize"]
    includes:
      in_header: "preamble.tex"
editor_options: 
  chunk_output_type: console
---
```{=latex}
\justifying
```
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
# Used for rendering the manuscript rmarkdown document
library("papaja")
# Used for general data wrangling
library(tidyverse)
# Used to make exact file paths
library(here)
# Used for tables
library(kableExtra)
library(gt)
# Used for joining figures
library(patchwork)
# The developer version is needed
# Used for creating Likert ggplots
library(likert)
# Used for knitting rmarkdown documents
library(knitr)
# Used for colorblind ggplot colors
library(viridis)
# References
r_refs("ecaw-ref.bib")
```

@@

\begin{tcolorbox}
{\fontsize{9pt}{9pt}\selectfont\textbf{Box 1. A hypothetical example of an ECAW in practice.}}

\includegraphics[width=1\textwidth]{./ecaw_workflow.png}
{\fontsize{9pt}{9pt}\selectfont
The overarching framework for ECAWs is depicted above. Within this framework, data management organizations would need to pre-specify several details for each step. Below, we provide a hypothetical example for illustrative purposes. The exact implementation of ECAWs would depend on the preferences of each data management organization and the community who uses their dataset.

@@

# Results
## Survey completion
Of 1148 emails sent, 1094 went through and 54 bounced. The survey was completed `r nrow(processed)` times and partially completed `r nrow(processed_all) - nrow(processed)` times, leading to a response rate of `r round(nrow(processed) / 1094 * 100)`% for complete surveys and `r round(nrow(processed_all) / 1094 * 100)`% for at least partially complete surveys.\footnote[1]{The ALSPAC mailing list has been maintained as a record of collaborators for many years and is constantly added to. The ALSPAC team only removes email addresses from their mailing list if someone explicitly requests this action. Thus, their list likely includes several email addresses that are no longer monitored. For example, we received one email reply stating that the recipient hasn’t been active in research for 30 years. It is also possible that some researchers have multiple email addresses on the mailing list (e.g., because they moved institutions). These two factors may have deflated the response rate.} The median time to complete the survey was `r format(round(median(processed$duration_in_mins), 1), nsmall = 1)` minutes (IQR: `r format(round(quantile(processed$duration_in_mins, .25, na.rm = TRUE), 1), nsmall = 1)` to `r format(round(quantile(processed$duration_in_mins, .75, na.rm = T), 1), nsmall = 1)`). This manuscript presents the results for complete surveys. Supplementary Material D presents the results with partially complete surveys included.

@@

Most respondents agreed that studies that analyze preexisting observational datasets are trustworthy (`r pull(filter(typically_trustworthy, support == "positive"), percentage)`%; `r pull(filter(typically_trustworthy, support == "positive"), n)`/`r pull(filter(typically_trustworthy, support == "positive"), n_sum)`) and reproducible (`r pull(filter(typically_reproducible, support == "positive"), percentage)`%; `r pull(filter(typically_reproducible, support == "positive"), n)`/`r pull(filter(typically_reproducible, support == "positive"), n_sum)`) (Figure 1, top panel). At the same time, many agreed that a study using an ECAW would be _more_ trustworthy (`r pull(filter(ecaw_trustworthy, support == "positive"), percentage)`%; `r pull(filter(ecaw_trustworthy, support == "positive"), n)`/`r pull(filter(ecaw_trustworthy, support == "positive"), n_sum)`) and _more_ reproducible (`r pull(filter(ecaw_reproducible, support == "positive"), percentage)`%; `r pull(filter(ecaw_reproducible, support == "positive"), n)`/`r pull(filter(ecaw_reproducible, support == "positive"), n_sum)`) compared to a typical study using preexisting observational data (Figure 1, bottom panel).

```{r typicallyEcawPlot, warning=FALSE, fig.cap="(ref:typicallyEcawPlotCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=12.8, fig.height=6, fig.path='figs/', dev=c('png', 'pdf')}
# Prepare plot data
typically_plot_data <-
  processed %>%
  select(contains("typically_")) %>%
  mutate(
    across(
      .fns = ~ factor(., levels = c(
        "Strongly disagree",
        "Somewhat disagree",
        "Neither agree nor disagree",
        "Somewhat agree",
        "Strongly agree"
      ))
    )
  ) %>%
  rename(
    Reproducible = typically_reproducible,
    Trustworthy = typically_trustworthy
  )

@@

```{=latex}
{\smaller[1] \singlespacing
```
(ref:typicallyEcawPlotCaption)  __Figure 1. Responses to the survey questions on trustworthiness and reproducibility of observational research with preexisting data and ECAWs.__ The survey defined trustworthy as “meaning that the results and conclusions of the publications are valid, reliable, rigorous, and accurate. That they merit trust”. The survey defined reproducible “in the sense that other researchers re-analysing the data with the same research question would produce similar results.” For each item, the number to the left of the data bar indicates the combined percentage for the responses depicted in any shade of brown/orange. The number in the center of the data bar (gray) indicates the percentage of neutral responses. The number to the right of the data bar indicates the combined percentage for the responses depicted in any shade of green. The bar charts in the top panel had no missing responses or selection of the option _“I don’t understand the question”_. The bar charts in the bottom panel excluded responses of _“I don't understand the question”_ (n = `r pull(filter(ecaw_missing, item == "Trustworthy"), n)`; `r pull(filter(ecaw_missing, item == "Reproducible"), n)`).
```{=latex}
}
```

@@

```{r methodPlot, warning=FALSE, fig.cap="(ref:methodPlotCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=11, fig.height=4.5, fig.path='figs/', dev=c('png', 'pdf')}
# Used methods figures
method_plot_data <-
  processed %>%
  select(starts_with("method_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "variable",
    values_to = "scale",
  ) %>%
  separate(variable, into = c("name_prefix", "type")) %>%
  group_by(type) %>%
  count(scale) %>%
  ungroup() %>%
  replace_na(list(scale = "Missing")) %>%
  mutate(
    scale = factor(scale, levels = c(
      "Never or almost never",
      "Sometimes",
      "About half the time",
      "Most of the time",
      "Always or almost always",
      "Missing",
      "I don't understand the question"
    )),
    type = case_when(
      type == "blind" ~ "Blind the data analyst",
      type == "confirmatory" ~ "Contain confirmatory analysis",
      type == "exploratory" ~ "Contain exploratory analysis",
      type == "preregistered" ~ "Are preregistered",
      type == "script" ~ "Share analysis scripts",
    ),
    type = factor(type, levels = c(
      "Contain exploratory analysis",
      "Contain confirmatory analysis",
      "Blind the data analyst",
      "Share analysis scripts",
      "Are preregistered"
    ))
  ) %>%
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>%
  mutate(
    percentage = n / sum(n)
  ) %>%
  ungroup()

@@

```{=latex}
{\smaller[1] \singlespacing
```
(ref:methodPlotCaption) __Figure 2. Responses to survey questions about the research practices of participants.__
```{=latex}
}
```

@@

```{r alspacPlot, warning=FALSE, message=FALSE, fig.cap="(ref:alspacPlotCaption)", fig.align="center", fig.pos="H", out.width="100%", fig.width=14, fig.height=5, fig.path='figs/', dev=c('png', 'pdf')}
# Check if there is any missing and don't understand responses separately
alspac_missing <-
  processed %>%
  select(contains("alspac_")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "item",
    values_to = "value"
  ) %>%
  dplyr::filter(is.na(value) | value == "I don't understand the question" | value == "Unsure") %>%
  mutate(
    item = str_to_title(str_remove(item, "^[^_]*_")),
    # We reorder it by making it a factor
    item = factor(item, levels = c(
      "Less_willing",
      "Opt_in",
      "Study",
      "Prefer_ecaw"
    )),
    # By making it a factor we can complete missing levels in the data later
    value = factor(value, levels = c(
      "Missing",
      "I don't understand the question",
      "Unsure"
    ))
  ) %>%
  replace_na(list(value = "Missing")) %>%
  group_by(item, value) %>%
  summarise(
    n = n()
  ) %>% 
  # Add n = 0 if no one choose a specific factor level
  tidyr::complete(value, fill = list(n = 0))

@@

```{=latex}
{\smaller[1] \singlespacing
```
(ref:alspacPlotCaption) __Figure 3. Responses to survey questions about using ECAWs.__ These bar charts exclude responses of _"I don't understand the question"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "I don't understand the question"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "I don't understand the question"), n)`; respectively from top to bottom), and responses of _"Unsure"_ (_n_ = `r pull(filter(alspac_missing, item == "Less_willing" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Opt_in" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Study" & value == "Unsure"), n)`; `r pull(filter(alspac_missing, item == "Prefer_ecaw" & value == "Unsure"), n)`). Agreement with the first question may be slightly inflated due to the format of the questions in this block. Respondents with a highly positive inclination towards ECAWs would be expected to disagree with the first question, but agree with the next three questions. Four respondents agreed with all four statements, suggesting they may have glazed over the word “less” in the first question \protect\footnotemark[3]. Interpreting responses to the second and third question come with a degree of ambiguity as the survey did not specify what was meant by the term “study” \protect\footnotemark[4].
```{=latex}
}
```

@@

```{=latex}
{\smaller[1]
\begin{singlespace}
```
___Table 1. Recurring topics in responses to the open-ended survey questions.___ The survey included 4 open-ended questions with broad prompts regarding running a study on ECAWs, benefits and drawbacks of ECAWs, related research practices, and general comments. These questions received a total of (`r nrow(open_ended)`) responses from (`r nrow(distinct(open_ended, response_id))`) unique respondents. A complete list of responses are viewable in the open data [LINK]. We synthesized the responses to open-ended questions into the 9 topics on the left side of this table. We divide these into three sections: (i) concerns about the acceptability of ECAWs, (ii) concerns that ECAWs will not have their intended impact, and (iii) alternative interventions that may achieve similar goals as conventional preregistration and ECAWs. On the right side of the table, we provide our reflection on each topic.
```{=latex}
\end{singlespace}
}
```

```{r openEndedTable, echo=FALSE, message=FALSE, warnings=FALSE, fig.align="center", out.width="100%", out.height="300%", fig.id=FALSE}

@@

open_ended_table %>% 
  group_by(subheader) %>% 
  gt::gt(.) %>% 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_row_groups(), cells_column_labels())
    ) %>% 
  tab_footnote(
    footnote = "The text in the Respondents’ comments column is synthesized from multiple responses and/or paraphrased. They are not verbatim responses.",
    locations = cells_column_labels(
      columns = `Respondents’ comments`
    )
  ) %>% 
  tab_footnote(
    footnote = "Sarafoglou et al.’s study used a pre-cleaned dataset. We are not aware of data that elucidates the time difference between ECAW-like workflows and conventional preregistration when starting from raw data.",
    locations = cells_body(
      columns = `Our reflection`,
      rows = 1
      )
  ) %>% 
  tab_options(
    table.width = pct(100)
    # table.font.size = 12
    ) %>% 
  gtsave("figs/open-ended-table_figure.png", expand = c(0, 50, 0, 50), vwidth = 1300, vheight=1500)

@@

## Exploratory analyses
Results can be explored interactively by following the instructions for running the code at [INSERT DATA LINK]. Based on the sample size of 107 participants and a lack of visually striking differences when exploring subsets of respondents, we do not report further on exploratory analyses.

@@

## Contributions
__Conceptualization:__ Robert T. Thibault and Marcus R. Munafò.\hfill\break
__Data curation:__ Marton Kovacs.\hfill\break
__Formal analysis:__ Robert T. Thibault and Marton Kovacs.\hfill\break
__Funding acquisition:__ Robert T. Thibault and John P. A. Ioannidis.\hfill\break
__Investigation:__ Robert T. Thibault.\hfill\break
__Methodology:__ Robert T. Thibault, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.\hfill\break
__Project administration:__ Robert T. Thibault.\hfill\break
__Resources:__ Robert T. Thibault.\hfill\break
__Software:__ Marton Kovacs.\hfill\break
__Supervision:__ Robert T. Thibault and Marcus R. Munafò.\hfill\break
__Validation:__ Robert T. Thibault and Marton Kovacs.\hfill\break
__Visualization:__ Robert T. Thibault and Marton Kovacs.\hfill\break
__Writing - original draft:__ Robert T. Thibault.\hfill\break
__Writing - review & editing:__ Robert T. Thibault, Marton Kovacs, Tom E. Hardwicke, Alexandra Sarafoglou, John P. A. Ioannidis, and Marcus R. Munafò.

@@
